ANGSD-wrapper: utilities for analyzing next generation sequencing data

### 2016-05-23 manuscript.tex
minor last edits

\begin{abstract}
High throughput sequencing has changed many aspects of population genetics, molecular ecology, and related fields, affecting both experimental design and data analysis. 
The software package ANGSD allows users to perform a number of population genetic analyses on high-throughput sequencing data.
ANGSD uses probabilistic approaches which can directly make use of genotype likelihoods, thus SNP calling is not a required for comparative analyses. This takes advantage of all of the sequencing data and producing more accurate results for samples with low sequencing depth.
Here we present ANGSD-wrapper, a set of wrapper scripts that provide a user-friendly interface for running ANGSD and visualizing results.
ANGSD-wrapper supports multiple types of analyses including estimates of nucleotide sequence diversity and performs neutrality tests, principal component analysis, estimation of admixture proportions for individuals samples, and calculation of statistics that quantify recent introgression. 
ANGSD-wrapper also provides interactive graphing of ANGSD results to enhance data exploration.
We demonstrate the usefulness of ANGSD-wrapper by analyzing resequencing data from populations of wild and domesticated \textit{Zea}. 
ANGSD-wrapper is freely available from \url{https://github.com/mojaveazure/angsd-wrapper}.
\end{abstract}

@@

A number of tools have recently been published to estimate population genetic descriptive statistics using high throughput sequencing data \citep{garrigan2013popbam, purcell2007plink, danecek2011variant, hutter2006genome}. The software package ANGSD \citet{korneliussen2014angsd}, is noteworthy because it enables users to perform a large number of common population genetic analyses, including estimation of diversity statistics, admixture analysis including Patterson's D statistic \citep{Durand:2011jd}, site frequency spectrum estimation \citep{pmid22911679}, and calculation of neutrality test statistics \citep{korneliussen_calculation_2013}. 
ANGSD works directly with alignment formats produced from standard high throughput sequence analysis pipelines, which removes the need for the user to transform the data into a software-specific format. 
One of the most important features of ANGSD is that analyses are integrated over per-site genotype likelihoods, rather than on predetermined polymorphic sites \emph{a priori}. 
This permits ANGSD to calculate common population genetic descriptive statistics on low-coverage sequencing data, and handle missingness due to variation in coverage. 
Further, this probabilistic framework is appealing to researchers studying non-model organisms where the lack of existing genomic resources or cost considerations prevent the ability to obtain high quality data. 
Also, many existing tools do not implement statistical models that account for biological issues such as non-random mating. Recent use of ANGSD highlights these benefits, allowing population genetic analyses in non-model organisms such as the blue-eyed black lemur \citep{meyer2015evolutionary} and \textit{Ficedula} flycatchers \citep{burri2015linked} and inbred lines of wild and cultivated maize \citep{Beissinger031666}.

>>>>

\begin{abstract}
High throughput sequencing has changed many aspects of population genetics, molecular ecology, and related fields, affecting both experimental design and data analysis. 
The software package ANGSD allows users to perform a number of population genetic analyses on high-throughput sequencing data.
ANGSD uses probabilistic approaches which can directly make use of genotype likelihoods, and SNP calling is thus not required for comparative analyses. This takes advantage of all of the sequencing data and produces more accurate results for samples with low sequencing depth.
Here we present ANGSD-wrapper, a set of wrapper scripts that provide a user-friendly interface for running ANGSD and visualizing results.
ANGSD-wrapper supports multiple types of analyses including estimates of nucleotide sequence diversity and performs neutrality tests, principal component analysis, estimation of admixture proportions for individuals samples, and calculation of statistics that quantify recent introgression. 
ANGSD-wrapper also provides interactive graphing of ANGSD results to enhance data exploration.
We demonstrate the usefulness of ANGSD-wrapper by analyzing resequencing data from populations of wild and domesticated \textit{Zea}. 
ANGSD-wrapper is freely available from \url{https://github.com/mojaveazure/angsd-wrapper}.
\end{abstract}

@@

A number of tools have recently been published to estimate population genetic descriptive statistics using high throughput sequencing data \citep{garrigan2013popbam, purcell2007plink, danecek2011variant, hutter2006genome}. The software package ANGSD \citet{korneliussen2014angsd}, is noteworthy because it enables users to perform a large number of common population genetic analyses, including estimation of diversity statistics, admixture analysis including Patterson's D statistic \citep{Durand:2011jd}, site frequency spectrum estimation \citep{pmid22911679}, and calculation of neutrality test statistics \citep{korneliussen_calculation_2013}. 
ANGSD works directly with alignment formats produced from standard high throughput sequence analysis pipelines, which removes the need for the user to transform the data into a software-specific format. 
One of the most important features of ANGSD is that analyses can be integrated over all sites, rather than on predetermined polymorphic sites \emph{a priori}. 
This permits ANGSD to calculate common population genetic descriptive statistics on low-coverage sequencing data, and handle missingness due to variation in coverage. 
Further, this probabilistic framework is appealing to researchers studying non-model organisms where the lack of existing genomic resources or cost considerations prevent the ability to obtain high quality data. 
Also, many existing tools do not implement statistical models that account for biological issues such as non-random mating. Recent use of ANGSD highlights these benefits, allowing population genetic analyses in non-model organisms such as the blue-eyed black lemur \citep{meyer2015evolutionary} and \textit{Ficedula} flycatchers \citep{burri2015linked} and inbred lines of wild and cultivated maize \citep{Beissinger031666}.
